---
title: "Wealth and Environmental Stewardship"
author: "Felipe Valencia"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Load Libraries
library(tidyverse)
library(downloader)
library(readxl)
library(stringr)
library(psych)
library(vtable)
```

```{r load_data}
# Import data

# Read xlsx
# Temporary file
temp_excel <- tempfile()
# Download using downloader package
download("https://datadryad.org/stash/downloads/file_stream/153034",
         dest = temp_excel, # Save temp file
         mode = "wb")
# Read xlxs from temp file
BEV_FCV <- read_excel(temp_excel)

```

## Background

__My question:__ Is it true that _as people get wealthier, they care and do more for the environment?_.

Here are the data sets I have found that I think will help me answer my question for this project:

+ [GDP per capita - current US dollars](https://data.worldbank.org/indicator/NY.GDP.PCAP.CD) from The World Bank.
    + This data set contains the GDP per capita in current US dollars for almost all countries with data from 1960 to 2021. According to Investopedia GDP per capita _"measures the economic output of a nation per person. It seeks to determine the prosperity of a nation by economic growth per person in that nation. Per capita income measures the amount of money earned per person in a nation."_
    + GDP per capita will be the socioeconomic metric.
+ [CO2 emissions (metric tons per capita)](https://data.worldbank.org/indicator/EN.ATM.CO2E.PC) from The World Bank.
    + This data set contains the CO2 emissions (metric tons per capita) for almost all countries with data from 1990 to 2019. The World Bank says: _"Carbon dioxide emissions are those stemming from the burning of fossil fuels and the manufacture of cement. They include carbon dioxide produced during consumption of solid, liquid, and gas fuels and gas flaring."_
    + CO2 emissions (metric tons per capita) will be our environmental metric.
+ [Encuestas de movilidad 2019](https://www.simur.gov.co/encuestas-de-movilidad) from Secretaría Distrital de Movilidad (SDM) of Bogota City Hall. (Bogota is the capital of Colombia)
    + This data set has information about travels, people, households, vehicles, etc. It's from a survey conducted by request of the Bogota City Hall to _"characterizes the mobility of the population equal to or greater than 5 years of age residing in the urban area of Bogotá and 18 neighboring municipalities in its area of influence: Bojacá, Cajicá, Chía, Cota, El Rosal, Facatativá, Funza, Gachancipá, La Calera, Madrid, Mosquera, Sibaté, Soacha, Sopó, Tabio, Tenjo, Tocancipá and Zipaquirá."_
    + This data set has the socioeconomic strata (to know the affluence) and types of vehicle (to see if there are EVs).
+ [Sociodemographic data for battery electric vehicle owning households in California (From NCST Project "Understanding the Early Adopters of Fuel Cell Vehicles")](https://datadryad.org/stash/dataset/doi:10.25338/B8P313) from Scott Hardman of the University of California and published in DRYAD.
    + This data set contains _"Sociodemographic data for fuel cell and battery electric vehicle owning households in California."_
    + This data set is from owners of BEVs and FCVs in California, it has their estimated income and a score measuring their perception of the importance of reducing greenhouse emissions.
+ [Number of passenger cars per adult in households, by household income. 2018-2019](https://www.ssb.no/en/transport-og-reiseliv/artikler-og-publikasjoner/the-wealthiest-bought-4-out-of-10-evs?tabell=417112) from the article by Statistics Norway called _"The wealthiest bought 4 out of 10 EVs"_.
    + This data set contains data from 2018 and 2019 about the number of passenger cars per adult in households by household income, it has data on fossil cars and electric cars.
    + This data set is already aggregated, therefore, there's not much to explore but it seems it can help solidify my answer.

## Battery Electric Vehicle (BEV)

This data set contains results from a survey of households in California. We are only going to focus on the: Information on vehicle owned, Household Income, Highest Level of Education, Age, Gender, Number of vehicles in the household, and a scale of the importance of reducing Greenhouse Gas Emissions (GGE). It contains around 906  Fuel-Cell Vehicles (FCV) respondants, but we are going to ignore them and focus on the BEVs respondents. We assume that those that own a BEV are _"environmentally friendly"_, and we are going to see if their wealth has an effect on that perspective.

### Data Wrangling

Here's the data we have cleaned and prepared for our analysis:

```{r tidy_data}
# Clean & wrangle data

BEV <- BEV_FCV %>%
  filter(!is.na(`submitdate. Date submitted`)) %>%
  rename(Household_Income = `Household Income`,
         ID = `id. Response ID`,
         importance_reduce_GGE = `Importance of reducing greenhouse gas emissions (-3 not important, 3 important)`,
         edu_level = `Highest Level of Education`,
         number_vehicles_H = `Number of vehicles in the household`,
         gender = `Gender (Male 1)`) %>%
  mutate(Household_Income_Thousands = Household_Income / 1000,
         edu_level = str_replace_all(as.character(edu_level), c("1" = "Some High School", "2" = "High School Graduate", "3" = "College Graduate", "4" = "Masters, Doctorate, \nor Professional Degree")),
         gender = str_replace_all(as.character(gender), c("1" = "Male", "0" = "Female"))) %>%
  select(ID, Carmain, Household_Income_Thousands, importance_reduce_GGE, edu_level, number_vehicles_H, gender, Age) %>%
  filter(grepl('PHASE', ID))

# Turn Household income in thousands, gender and edu_level to factors
BEV$Household_Income_Thousands <- as.factor(BEV$Household_Income_Thousands)
BEV$gender <- as.factor(BEV$gender)
BEV$edu_level <- as.factor(BEV$edu_level)
# Custom order for the edu_level factor
BEV$edu_level <- factor(BEV$edu_level, levels = c("Some High School", "High School Graduate", "College Graduate", "Masters, Doctorate, \nor Professional Degree"))
BEV

```
Here's a summary statistics table to understand the data we are working with:

```{r stats1}
# Get Summary Statistics
BEV %>% st()

```

### Data Visualization

```{r plot_data1}
# Plot & visualize data
BEV %>% ggplot(aes(x = importance_reduce_GGE)) +
  geom_histogram(binwidth = 0.3, color = "black", fill = "green4") +
  theme_classic() +
  scale_x_continuous(breaks = c(-3, -2, -1, 0, 1, 2, 3),
                     expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  theme(panel.grid.major.y = element_line(color = "gray", linetype = 1),
        plot.title.position = "plot",
        plot.title = element_text(size = 25, family = "serif", color = "black"),
        plot.subtitle = element_text(size = 20, family = "serif", color = "gray30"),
        plot.caption = element_text(hjust = 1, family = "serif", color = "gray30", size = 12),
        axis.title = element_text(color = "gray30", size = 18, family = "serif"),
        axis.text.x = element_text(color = "gray30", size = 15, family = "serif"),
        axis.text.y = element_text(color = "gray30", size = 15, family = "serif")) +
  labs(title = "Distribution of attitudes towards sustainability", subtitle = "Respondents were asked about how important is reducing greenhouse gas emissions. \nMeasured with a continuous scale from -3= “Not important” to 3= “Important”", y = "Count", x = "Scale", caption = "Source: National Center for Sustainable Transportation\nhttps://doi.org/10.25338/B8P313")

```
Going back to the summary statistics table, the mean is around __1.7__, which tell us as that there's an attitude of care for the sustainability of the environment from owners of BEVs. With the histogram we can see the majority of the respondents consider it's important to reduce greenhouse gas emissions.

Now, let's focus on the wealth of the owners of BEVs in California. According to the summary statistics, we have the majority of the respondents with an income between _$75k_ and _$225k_, but we have enough data for incomes greater than _$225k_ to see if on average people with higher income care more about the environment than those with lower incomes among those that we consider _"environmentally friendly"_ due to owning a BEV.

```{r plot_data2}
# Plot & visualize data

# Turn Household income in thousands to a numeric data type
BEV$Household_Income_Thousands <- as.numeric(as.character(BEV$Household_Income_Thousands))

BEV %>%
  filter(!is.na(Household_Income_Thousands)) %>%
  filter(!is.na(edu_level)) %>%
  ggplot(aes(y = importance_reduce_GGE, x = Household_Income_Thousands)) +
  geom_smooth(color = "green4") +
  theme_bw()
  #coord_cartesian(ylim = c(-3, 3), xlim = c(50, 500))
```

```{r plot_data3}
# Plot & visualize data
BEV %>%
  filter(!is.na(Household_Income_Thousands)) %>%
  filter(!is.na(edu_level)) %>%
  ggplot(aes(y = importance_reduce_GGE, x = Household_Income_Thousands)) +
  geom_boxplot(aes(group = Household_Income_Thousands)) +
  geom_smooth(color = "green4") +
  facet_wrap(~edu_level, nrow = 1) +
  theme_bw()
  #coord_cartesian(ylim = c(1.2, 3), xlim = c(50, 500))
```

```{r plot_data4}
# Plot & visualize data

```
### Conclusions

### Discussion

The only limitation with this data set is that we don't have information about the attitude of non-BEVs owners, like owners of fuel-based vehicles to compare if there are differences in their perspectives and their income.
